Project Description -- Erosion, Transport, Anomalous Diffusion, and Collapse in Granular and Porous Media

### 2019-11-17 docs/proposals/NSF/ProjDesc/ProjDesc.tex
worked through proposal

  \item Simulate transport of finite Peclet number species in eroded
    geometries.
\end{itemize}

@@

An immediate goal of the proposed work is to extend the range of physical effects captured in these simulations. A particular challenge of a BIE formulation is developing quadrature for nearly-singular integrands that arise when bodies are close. To address this problem, we have very recently extended a Barycentric quadrature~\cite{bar2014, bar-wu-vee2015} and have successfully simulated erosion with grains that are initially separated by 1\% of an arclength spacing~\cite{chi-moo-qua2019}. This quadrature approach has the advantage that it is a non-intrusive modification of the trapezoid rule, allowing us to achieve an error bound that independent of the grain configuration. However, the implementation of the quadrature method does not fully use FMM acceleration. The numerical challenge is developing the required FMM expansions for the velocity, vorticity, and deformation tensor layer potentials. Since we require several layer potentials with different kernels, we will utilize kernel-independent FMMs~\cite{yin-bir-zor2004, fon-dar2009}, and its implementations that are readily available (for example, see \url{http://pvfmm.org}). Other challenges that will be investigated include preconditioning strategies to reduce the number of required GMRES iterations, and time adaptivity to allow for stable simulations to user-specified tolerances. PI Quaife has experience developing such algorithms~\cite{qua-bir2015b, qua-bir2016, qua-bir2015a, qua-cou-dar2018}.

@@

Transport through eroded media is a direct consequence of the pore structure. Qualitative properties of an eroded media, such as channels of high porosity, can be modest in volume fraction, but can be responsible for a large portion of the flux~\cite{Quaife2018}.  To increase understanding of flow through eroded media, we will simulate and statistically analyze transport through eroded geometries.

In an eroded geometry $\Omega$, the dimensionless
concentration $c$ is governed by
\begin{align}
  \pd{c}{t} + \uu \cdot \nabla c = \frac{1}{\Pe} \Delta c, 
  \label{eqn:advectionDiffusion}
\end{align}
where $\Pe$ is the dimensionless Peclet number. Here we assume that the
concentration $c$ is affected by the eroded geometry, but the geometry
is not altered further by the concentration. In the infinite Peclet
number limit, the concentration is governed by pure advection and the
characteristics of the concentration field are identical to the
streamlines (Fig.~\ref{fig:100tracers}).

The $j^{th}$ trajectory, or characteristic, is governed by $\dot{\ss}_j(t) = \uu(\ss_j(t))$. Once these trajectories are computed, using for example a high-order Runge-Kutta method, a statistical analysis can be used to describe the transport. Two of the most important measurements are the tortuosity and the anomalous dispersion, both which are defined in terms of the length of the trajectories \begin{align}
  \lambda_j(t) = \int_{0}^{t} \|\ss'(t')\| dt'.
\end{align}
Upon computing $N_p$ trajectories, the tortuosity is the dimensionless number
\begin{align}
  T = \frac{1}{d} \left(\int_{S} u_1(x_0,y_0)\lambda(y_0) dy_0\, \right)
  \Bigg/ \left(\int_S u_1(x_0,y_0)\, dy_0 \right),
\end{align}
where $S$ is the vertical cross-section at $x=x_0$, $d$ is its length,
$u_1$ is the horizontal velocity, and $\lambda(y_0)$ is the length of
the trajectory starting at $(x_0,y_0)$ and ending at a vertical
cross-section at the other end of the channel. Relationships between the
tortuosity and porosity of a geometry have been developed by other
groups~\cite{kop-kat-tim1996, dud-koz-mat2011, mat-kha-koz2008}, and
preliminary results indicate that such models might be appropriate for
geometries whose porosity is increasing under the process of erosion
(Fig.~\ref{fig:100tortuosity}(a)). However, preliminary results also
indicate that erosion can increase the amount of tortuosity as is
evident towards the at the towards the end of the erosion process in
Fig.~\ref{fig:100tortuosity}(a). Moreover, we have observed a
significant increase in the tortuosity at small porosities
(Fig.~\ref{fig:100tortuosity}(b)).

@@

While the tortuosity characterizes the amount of winding of streamlines, the dispersion characterizes the spreading of the trajectories.  Dispersion can be described using the first- and second-ensemble moments of the trajectories
\begin{align}
  \langle \lambda \rangle (t) = \frac{1}{N_p} \sum_{j=1}^{N_p}
    \lambda_j(t), \quad \sigma_\lambda^{2}(t) = \frac{1}{N_p}
    \sum_{j=1}^{N_p} (\lambda_j(t) - \langle \lambda \rangle(t))^2
\end{align}
Then, the standard deviation $\sigma_\lambda$ characterizes the
dispersion, and long-time behavior typically results in a power-law
scaling $\sigma_\lambda \sim t^{\alpha}$. The power $\alpha$, which
quantifies the long-time spreading, is $1$ in a ballistic regime which
occurs at early times when the trajectories have not explored much of
the porous geometry. However, once the trajectories have traversed a few
grains, dispersion in porous media is often super-dispersive with
$\alpha \in (0.5,1)$~\cite{dea-leb-den-tar-bol-dav2013}. In Fig.~\ref{fig:100dispersion}, we investigate spreading at several different porosities of a porous geometry that is initialized with 100 grains.
\begin{wrapfigure}[22]{r}{0.4\textwidth}
  \includegraphics[width=0.4\textwidth]{figs/100b_second_moment_long_ref}
  \caption{\label{fig:100dispersion} The particle spreading,
  characterized with the standard deviation $\sigma_\lambda(t)$ at 6
  different porosities. At early times, the line of slope 1 (black
  curve) indicates a ballistic motion. At later times, the
  dispersion becomes super-dispersive at a rate indicated by the slope
  of the black dashed-dotted line.}
\end{wrapfigure}

@@

Near-term plans are to include flow transport and gravitational sedimentation of the eroding grains. For this extension, contact forces between bodies must be included to eliminate unphysical contact that is inevitable due to discretization errors.

Though conceptually straightforward, implementing contact forces in multibody fluid-structure simulations is far from simple in practice, with several competing possible strategies that are a topic of current research.  One approach is to use introduce a potential with short-range repulsion, but such methods often introduce stiffness. Alternatively, a space-time interference volume can guarantee a minimum separation between grains.  In a fluid context, this method was first introduced by Lu et al.~\cite{lu-rah-zor2017}, and extended to allow for much smaller separation distances of rigid bodies by the PI's previous graduate student~\cite{bys-sha-qua2019}. One PhD student, who will be jointly supervised by both PIs, will simulate sedimentation of eroding bodies by developing algorithms that resolve quadrature error and contact.

% PREVIOUS PARAGRAPH
%Near-term plans are to include flow transport and gravitational sedimentation of the particles. For this extension, contact forces between bodies must be included. Though conceptually straightforward, implementing contact forces in multibody fluid-structure simulations is far from simple in practice, with several competing possible strategies that are a topic of current research  {\color{blue} cite}. We plan to implement contact forces with {\color{blue} stuff}. This goal will form the PhD dissertation of a graduate student who will be supervised by myself and Prof.~Bryan Quaife (FSU Scientific Computing).

One central goal of this proposal is to enable the free motion of grains during the erosion processes. As mentioned earlier, this goal requires resolving contact forces between bodies, as well as the incorporation of the external buoyancy forces. Once this goal is achieved, it will be possible to simulate a complex porous-medium of arbitrarily sized and shaped grains undergoing the simultaneous action of mechanical-or-chemical erosion and transport due to the combined effects of seepage, contact, and buoyancy forces. These computations will thus simulate, with high-fidelity, the multi-physics of realistic porous-medium dynamics, albeit in a scaled down context. Ultimately, such simulations could be used to build an understanding of medium reconfiguration events, such as sinkhole collapse or conduit formation in karst networks.

In very preliminary work, PI Moore has been running simplified simulations with a current graduate student. These preliminary simulations are based on a hybrid continuum-discrete model for the permeability field and grain particle field respectively. They utilize the homogenized description of the flow field in~\eqref{DarcyEq} and incorporate the physical effects of:
\begin{itemize}[noitemsep]
\item Transport of grains due to seepage force from the surrounding fluid flow.
\item Gravitational sedimentation of grains.
\item Feedback of grain-distribution onto permeability field.
\item Contact and cohesive forces between grains.
\end{itemize}
Preliminary investigation suggests that this last mechanisms of cohesion is essential to model sinkhole collapse in the system.

\todo[inline]{Eventually, our FMM solver will reach its capacity. We
need to talk about fast summation methods for the barycentric
quadrature, deformation tensor, vorticity, and pressure. In addition to
fast summation methods, we can talk about parallel implementations. I
have new PhD student in DSC that wants to work on erosion and he has
experience running MPI and OpenMP.}

These simulations have not yet begun to incorporate the slow-acting effects of mechanical and chemical erosion, and the associated gradual changes in medium properties. A grand goal of this proposal would be to merge the two lines of inquiry to develop a sound theoretical framework for porous-medium dynamics robust enough to capture both short-term events, such as sinkhole collapse, and long-term changes in the medium. The greatest challenge faced is the disparate temporal and spatial scales that must be considered. Our proposed strategy is to use the high-fidelity erosion simulations to parametrize the effect of erosion on medium properties.

%----------------------------------------------------------------------%
\subsection{Parameterizations of Eroding Porous Media}
More specifically, we plan to run large batches of the high-fidelity boundary-integral-equation-based numerical simulations of erosion and then leverage rapid advances in the field of machine learning and deep-neural networks to develop reduced models that parametrize the effects of erosion on the evolution of macroscopic medium properties and the flow.  The output of the high-fidelity simulations of the grain configuration can be transformed into essentially an image format, enabling us to use cutting-edge advancements on deep neural-networks for image analysis~\cite{pelt2018mixed}. Some important advantages of referenced work are the use of dilated convolutions in a deeply connected network, which first mixes scales and second reduces the number of super-parameters present and thus largely averts the risk of overfitting data. Additionally, high-fidelity flow conditions can be used to train a neural network, and this has been successfully used by others to predict, as one example, an anisotropy tensor~\cite{ling2016reynolds}.  Both PIs will supervise a graduate student who will design and implement the deep-neural network for learning the essence of the high-fidelity simulations. Once the essential features are learned, they will be incorporated into reduced-order models that utilize the multiphase framework to enable more efficient and larger scale coarse-grained simulations of dynamic erosion.

\subsection{Experimental Validation}
\label{sec:experiments}
As mentioned previously, researchers at FSU's Geophysical Fluid Dynamics Institute (GFDI) have investigated the mechanisms behind sinkhole formation with a series of scaled-down laboratory experiments~\cite{tao2014experimental}. First, already existing data from these experiments offers a testbed for comparison against our newly developed simulations. Furthermore, we propose to run a set of new experiments at the facilities offered by GFDI, spearheaded by an undergraduate student in close collaboration with the graduate students, PIs, and an experimentalist at GFDI. These new experiments will more deeply probe the mechanisms underlying sinkhole formation with the special design goal of comparison against the numerical simulations. These experiments will be used to calibrate and validate our computations.

%----------------------------------------------------------------------%
%%%%  SECTION: Education and outreach
%----------------------------------------------------------------------%
\section{Outreach}
The PIs are committed to ....

%----------------------------------------------------------------------%
\subsection{Graduate research}
An important benefit of the research proposed here is the potential to immerse graduate students in fundamentally {\em cross-disciplinary} research. This proposal involves a confluence of disciplines, most centrally computational mathematics, but also geophysics, reduced-physical modeling, hydrology, and environmental science.  Exposure to a range of disciplines is, in the view of the PIs, an invaluable opportunity for young scientists.

@@

%----------------------------------------------------------------------%
\subsection{Undergraduate research}
Regarding undergraduate research, the ideas in this proposal represent a rich and fascinating topic for undergraduate students, while also being approachable these students. In particular, we aim to recruit at least one undergraduate student to perform laboratory experiments at the Geophysical Fluid Dynamics Institute (GFDI). As mentioned above, GFDI offers state-of-the-art laboratory facilities, with an already-functioning sinkhole experiment. The student will perform controlled experiments using natural and artificial porous media (e.g.~sand and micro-glass beads respectively) to precisely characterize the parameter regime leading to gravitational collapse. Parallel experiments will be conducted with erodible (bentonite clay) or dissolvable  (sucrose solids) fragments materials, which will represent the porous medium. These experiments will be used to guide, calibrate, and validate the computational advances. These undergraduate students will work closely with the graduate students to ensure that theory and experiments advance in close cooperation, with constant feedback between the two.

@@

%----------------------------------------------------------------------%
\subsection{Outreach for K--12 and the general public}
The PIs have already seized the opportunity to engage K--12 students in outreach activities. In particular, PI Moore's research was the focus of an educational video created by {\bf CPALMS}, which is the State of Florida's official source for standards information and course descriptions for K--12 education. The video, which features an interview and animations of research, is used to reinforce concepts from mathematics courses (grade levels 7--12), and to encourage students to consider a career in the STEM fields~\cite{CPALMS}. Additionally, these videos are frequently used for the continuing education of K--12 {\em teachers}, providing them a broader perspective of mathematics that could be integrated into their teaching.

@@

%\subsection{Assessment for graduate and undergraduate education}
%Developing strong mentoring relationships is essential to a student's intellectual development and a responsibility that the PIs takes very seriously. When working with students on research, whether undergraduate or graduate, we are always sure to outline the expectations of {\em both} parties -- the student and myself -- and periodically revisit the list to determine if the expectations are being met or if changes are necessary. This ensures that progress continues uninterrupted and that both parties are comfortable with their responsibilities. For the graduate seminar course, questionnaires will be provided to the students to determine how the seminar relates to their own research and their intellectual development, as well as parts of the course that were particularly beneficial and parts that could be improved. This will ensure that, once integrated into the standard curriculum, the course will maximally benefit the students' education.

>>>>

  \item Simulate transport of finite Peclet number species in eroded
    geometries.

  \item Understand pore size growth under erosion

  \item New parameterizations using modern machine learning techniques

  \item Validation with experimental results
\end{itemize}

@@

An immediate goal of the proposed work is to extend the range of physical effects captured in these simulations. A particular challenge of a BIE formulation is developing quadrature for nearly-singular integrands that arise when bodies are close. To address this problem, we have very recently extended a Barycentric quadrature~\cite{bar2014, bar-wu-vee2015} and have successfully simulated erosion with grains that are initially separated by 1\% of an arclength spacing~\cite{chi-moo-qua2019}. This quadrature approach has the advantage that it is a non-intrusive modification of the trapezoid rule, allowing us to achieve an error bound that independent of the grain configuration. However, the implementation of the quadrature method does not fully use FMM acceleration. The numerical challenge is developing the required FMM expansions for the velocity, vorticity, and deformation tensor layer potentials. Since we require several layer potentials with different kernels, we will utilize kernel-independent FMMs~\cite{yin-bir-zor2004, fon-dar2009}, and its HPC implementations that are readily available (for example, see \url{http://pvfmm.org}). Other challenges that will be investigated include preconditioning strategies to reduce the number of required GMRES iterations, and time adaptivity to allow for stable simulations to user-specified tolerances. PI Quaife has experience developing such algorithms~\cite{qua-bir2015b, qua-bir2016, qua-bir2015a, qua-cou-dar2018}.

@@

Transport through eroded media is a direct consequence of the pore structure. Qualitative properties of an eroded media, such as channels of high porosity, can be modest in volume fraction, but can be responsible for a large portion of the flux~\cite{Quaife2018}.  To increase understanding of flow through eroded media, we will simulate and statistically analyze transport through eroded geometries. In an eroded geometry $\Omega$, the dimensionless concentration $c$ is governed by
\begin{align}
  \pd{c}{t} + \uu \cdot \nabla c = \frac{1}{\Pe} \Delta c, 
  \label{eqn:advectionDiffusion}
\end{align}
where $\Pe$ is the dimensionless Peclet number. Here we assume that the concentration $c$ is affected by the eroded geometry, but the geometry is not altered further by the concentration (i.e.~$c$ is not responsible for dissolution).

In the infinite Peclet number limit, the concentration is governed by pure advection and the characteristics of the concentration field are identical to the streamlines (Fig.~\ref{fig:100tracers}). The $j^{th}$ trajectory, or characteristic, is governed by $\dot{\ss}_j(t) = \uu(\ss_j(t))$. Once these trajectories are computed, using for example a high-order Runge-Kutta method, a statistical analysis can be used to describe the transport. Two of the most important measurements are the tortuosity and the anomalous dispersion, both which are defined in terms of the length of the trajectories
\begin{align} 
  \lambda_j(t) = \int_{0}^{t} \|\ss'(t')\| dt'.
\end{align}
Upon computing $N_p$ trajectories, the tortuosity is the dimensionless number
\begin{align}
  T = \frac{1}{d} \left(\int_{S} u_1(x_0,y_0)\lambda(y_0) dy_0\, \right)
  \Bigg/ \left(\int_S u_1(x_0,y_0)\, dy_0 \right),
\end{align}
where $S$ is the vertical cross-section at $x=x_0$, $d$ is its length,
$u_1$ is the horizontal velocity, and $\lambda(y_0)$ is the length of
the trajectory starting at $(x_0,y_0)$ and ending at a vertical
cross-section at the other end of the channel. Relationships between the
tortuosity and porosity of a geometry have been developed by other
groups~\cite{kop-kat-tim1996, dud-koz-mat2011, mat-kha-koz2008}, and
preliminary results indicate that such models might be appropriate for
geometries whose porosity is increasing under the process of erosion
(Fig.~\ref{fig:100tortuosity}(a)). However, preliminary results also
indicate that erosion can increase the amount of tortuosity as is
evident towards the at the towards the end of the erosion process in
Fig.~\ref{fig:100tortuosity}(a). Moreover, we have observed a
significant increase in the tortuosity at small porosities
(Fig.~\ref{fig:100tortuosity}(b)).

@@

\begin{wrapfigure}[19]{r}{0.35\textwidth}
  \includegraphics[width=0.35\textwidth]{figs/100b_second_moment_long_ref}
  \caption{\label{fig:100dispersion} The particle spreading, characterized with the standard deviation $\sigma_\lambda(t)$ at six different porosities. At early times, the line of slope 1 (black curve) indicates a ballistic motion. At later times, the dispersion becomes super-dispersive at a rate indicated by the slope of the black dashed-dotted line.}
\end{wrapfigure}
While the tortuosity characterizes the amount of winding of streamlines, the dispersion characterizes the spreading of the trajectories.  Dispersion can be described using the first- and second-ensemble moments of the trajectories
\begin{align}
  \begin{split}
  \langle \lambda \rangle (t) &= \frac{1}{N_p} 
      \sum_{j=1}^{N_p} \lambda_j(t), \\ 
  \sigma_\lambda^{2}(t) &= \frac{1}{N_p}
    \sum_{j=1}^{N_p} (\lambda_j(t) - \langle \lambda \rangle(t))^2.
  \end{split}
\end{align}
Then, the standard deviation $\sigma_\lambda$ characterizes the
dispersion, and long-time behavior typically results in a power-law
scaling $\sigma_\lambda \sim t^{\alpha}$. The power $\alpha$, which
quantifies the long-time spreading, is $1$ in a ballistic regime which
occurs at early times when the trajectories have not explored much of
the porous geometry. However, once the trajectories have traversed a few
grains, dispersion in porous media is often super-dispersive with
$\alpha \in (0.5,1)$~\cite{dea-leb-den-tar-bol-dav2013}. In Fig.~\ref{fig:100dispersion}, we investigate spreading at several different porosities of a porous geometry that is initialized with 100 grains.

@@

Upon developing and improving numerical methods to simulate mechanical
and chemical erosion, we plan to develop methods to simulate transport,
gravitational sedimentation, and external buoyant forces of the eroding grains. In addition to the new previously described quadrature, preconditioning, adaptivity, and fast summation methods,  this extension also requires contact forces between bodies to eliminate unphysical contact that is inevitable due to discretization errors. Though conceptually straightforward, implementing contact forces in multibody fluid-structure simulations is far from simple in practice, with several competing possible strategies that are a topic of current research.  One approach is to use introduce a potential with short-range repulsion, but such methods often introduce stiffness. Alternatively, a space-time interference volume can guarantee a minimum separation between grains.  In a fluid context, this method was first introduced by Lu et al.~\cite{lu-rah-zor2017}, and extended to allow for much smaller separation distances of rigid bodies by the PI's previous graduate student~\cite{bys-sha-qua2019}. One PhD student, who will be jointly supervised by both PIs, will simulate sedimentation of eroding bodies by developing algorithms that resolve quadrature error and contact.

The central goal of simulating transport of eroding grains is to simulate a complex porous-medium of arbitrarily sized and shaped grains undergoing the simultaneous action of mechanical-or-chemical erosion and transport due to the combined effects of seepage, contact, and buoyancy forces. These computations will thus simulate, with high-fidelity, the multi-physics of realistic porous-medium dynamics, albeit in a scaled down context. Ultimately, such simulations could be used to build an understanding of medium reconfiguration events, such as sinkhole collapse or conduit formation in karst networks. In very preliminary work, PI Moore has been running simplified simulations with a current graduate student. These preliminary simulations are based on a hybrid continuum-discrete model for the permeability field and grain particle field respectively. They utilize the homogenized description of the flow field in~\eqref{DarcyEq} and incorporate the physical effects of: \begin{itemize}[noitemsep]
\item Transport of grains due to seepage force from the surrounding fluid flow.
\item Gravitational sedimentation of grains.
\item Feedback of grain-distribution onto permeability field.
\item Contact and cohesive forces between grains.
\end{itemize}
Preliminary investigation suggests that this last mechanisms of cohesion is essential to model sinkhole collapse in the system.

%----------------------------------------------------------------------%
\subsection{Parameterizations of Eroding Porous Media}
These simulations have not yet begun to incorporate the slow-acting effects of mechanical and chemical erosion, and the associated gradual changes in medium properties. A grand goal of this proposal would be to merge the two lines of inquiry to develop a sound theoretical framework for porous-medium dynamics robust enough to capture both short-term events, such as sinkhole collapse, and long-term changes in the medium. The greatest challenge faced is the disparate temporal and spatial scales that must be considered. Our proposed strategy is to use the high-fidelity erosion simulations to parametrize the effect of erosion on medium properties.

More specifically, we plan to run large batches of the high-fidelity boundary-integral-equation-based numerical simulations of erosion and then leverage rapid advances in the field of machine learning and deep-neural networks to develop reduced models that parametrize the effects of erosion on the evolution of macroscopic medium properties and the flow.  The output of the high-fidelity simulations of the grain configuration can be transformed into essentially an image format, enabling us to use cutting-edge advancements on deep neural-networks for image analysis~\cite{pelt2018mixed}. Some important advantages of referenced work are the use of dilated convolutions in a deeply connected network, which first mixes scales and second reduces the number of super-parameters present and thus largely averts the risk of overfitting data. Additionally, high-fidelity flow conditions can be used to train a neural network, and this has been successfully used by others to predict, as one example, an anisotropy tensor~\cite{ling2016reynolds}.  Both PIs will supervise a graduate student who will design and implement the deep-neural network for learning the essence of the high-fidelity simulations. Once the essential features are learned, they will be incorporated into reduced-order models that utilize the multiphase framework to enable more efficient and larger scale coarse-grained simulations of dynamic erosion.

%----------------------------------------------------------------------%
\subsection{Experimental Validation}
\label{sec:experiments}
As mentioned previously, researchers at FSU's Geophysical Fluid Dynamics Institute (GFDI) have investigated the mechanisms behind sinkhole formation with a series of scaled-down laboratory experiments~\cite{tao2014experimental}. First, already existing data from these experiments offers a testbed for comparison against our newly developed simulations. Furthermore, we propose to run a set of new experiments at the facilities offered by GFDI, spearheaded by an undergraduate student in close collaboration with the graduate students, PIs, and an experimentalist at GFDI. These new experiments will more deeply probe the mechanisms underlying sinkhole formation with the special design goal of comparison against the numerical simulations. These experiments will be used to calibrate and validate our computations.

%----------------------------------------------------------------------%
%%%%  SECTION: Education and outreach
%----------------------------------------------------------------------%
\section{Outreach}
The PIs are committed to utilizing and training students at various levels of their education throughout the proposed research. Different aspects of the proposed research are at appropriate levels for PhD students, undergraduate students, and high school students. Moreover, different opportunities within FSU will be used to publicize the work to K--12 students and the greater public.

%----------------------------------------------------------------------%
\subsection{Graduate Research}
An important benefit of the research proposed here is the potential to immerse graduate students in fundamentally {\em cross-disciplinary} research. This proposal involves a confluence of disciplines, most centrally computational mathematics, but also geophysics, reduced-physical modeling, hydrology, and environmental science.  Exposure to a range of disciplines is, in the view of the PIs, an invaluable opportunity for young scientists.

@@

%----------------------------------------------------------------------%
\subsection{Undergraduate Research}
Regarding undergraduate research, the ideas in this proposal represent a rich and fascinating topic for undergraduate students, while also being approachable these students. In particular, we aim to recruit at least one undergraduate student to perform laboratory experiments at the Geophysical Fluid Dynamics Institute (GFDI). As mentioned above, GFDI offers state-of-the-art laboratory facilities, with an already-functioning sinkhole experiment. The student will perform controlled experiments using natural and artificial porous media (e.g.~sand and micro-glass beads respectively) to precisely characterize the parameter regime leading to gravitational collapse. Parallel experiments will be conducted with erodible (bentonite clay) or dissolvable  (sucrose solids) fragments materials, which will represent the porous medium. These experiments will be used to guide, calibrate, and validate the computational advances. These undergraduate students will work closely with the graduate students to ensure that theory and experiments advance in close cooperation, with constant feedback between the two.

@@

%----------------------------------------------------------------------%
\subsection{Outreach for K--12 and the General Public}
The PIs have already seized the opportunity to engage K--12 students in outreach activities. In particular, PI Moore's research was the focus of an educational video created by {\bf CPALMS}, which is the State of Florida's official source for standards information and course descriptions for K--12 education. The video, which features an interview and animations of research, is used to reinforce concepts from mathematics courses (grade levels 7--12), and to encourage students to consider a career in the STEM fields~\cite{CPALMS}. Additionally, these videos are frequently used for the continuing education of K--12 {\em teachers}, providing them a broader perspective of mathematics that could be integrated into their teaching.

@@

%----------------------------------------------------------------------%
\section{Timeline and Concluding Remarks}
